Product plots

### 2011-03-07 prodplots-infovis.tex
Paper in infovis format

\documentclass[letterpaper,oneside]{scrartcl}
\usepackage{fullpage}
\usepackage[utf8]{inputenc}
\usepackage[pdftex]{graphicx}
\usepackage{tikz}
\DeclareGraphicsExtensions{.png,.pdf}
\graphicspath{{plots/}}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{url}
\usepackage[round,sectionbib]{natbib}
\bibliographystyle{abbrvnat}
\usepackage[small]{caption}
\usepackage[small]{titlesec}
\renewcommand\familydefault{bch}

\definecolor{female}{HTML}{EF8A62}
\definecolor{male}{HTML}{67A9CF}
\definecolor{not-too-happy}{HTML}{FC8D59}
\definecolor{pretty-happy}{HTML}{FFFFBF}
\definecolor{very-happy}{HTML}{91CF60}
\definecolor{grey30}{HTML}{4D4D4D}
\newcommand{\key}[1]
  {\protect \tikz{\fill[#1] rectangle (1ex,1ex);}}

\title{Product graphics}
\author{Hadley Wickham and Heike Hofmann}
\begin{document}
\maketitle

\section{Introduction}

% JASA methods
% Infovis?

% Aims of this paper:
%
%  * emphasise the connection between area graphics and factorising
%    probabilities
%
%  * pull all area plots on a common footing - provide a grammar for
%    categorical graphics

Area plots, like bar charts, mosaic plots and treemaps, are the graphical equivalent of contingency tables, visualising a discrete distribution by mapping proportion to area. This paper presents a unifying framework for area plots, showing how many existing plots are interrelated, and developing a set of tools for creating new plots. In particular, we create a graphical factorisation of area plots by analogy to the mathematical factorisation of joint distributions into products of marginals and conditionals.

@@

\begin{table}[htb]
  \begin{center}
  \begin{tabular}{llp{9cm}}
    \toprule
    Variable & Description & Values \\
    \midrule
    {\sf age} & age in years & 18--89 \\
    {\sf degree} & highest education & lt high school, high school, junior college, bachelor, graduate \\
    {\sf finrela} & relative financial status & far above average, above average, average, below average, far below average \\
    {\sf happy} & happiness & \key{very-happy} very happy, \key{pretty-happy} pretty happy, \key{not-too-happy} not too happy \\
    {\sf health} & health & excellent, good, fair, poor \\
    {\sf sex} & sex & \key{female} female, \key{male} male\\
    {\sf wtsall} & probability weight & 0.43--6.42 \\
    {\sf year} & year of survey & 1972--2006 \\
    \bottomrule
  \end{tabular}
  \end{center}
  \caption{Description of sample data. Common plot colours are shown next to respective levels.}
  \label{tbl:happy}
\end{table}

@@

\section{Modelling}
\label{sec:models}

\noindent There are 24 possible ways of ordering the variables in the pdf, and each 1d pmf can be displayed in five different ways. For a 4d problem this gives $5 * 24 * (5^3 + 5^4 + 5^2 + 1 + 5^2) = 96120$ plots!  How can we use statistical models to guide us towards an appropriate choice?

% Connection to models: \citep{theus:1999,hofmann:2001}

@@

@@

% bibtool -x product-graphics -o references.bib
\bibliography{references}
\end{document}

>>>>

% zip -r graphical-inference.zip *.pdf *.tex images references.bib *.cls

% 8 pages, including reference 
% 
% Frame as design paper:
% 
%   * explain the target problem
%   * enough background that the reader can pass judgement about whether your
%     solution is good
%   * you should present results that back up the claim that your approach is
%     better than others
%
% Supplemental material to create plots
\documentclass[journal]{vgtc}
\usepackage{mathptmx}
\usepackage{graphicx}
\graphicspath{{plots/}}
\usepackage{times}
\usepackage{booktabs}
\usepackage[utf8]{inputenc}
\usepackage[square,sort&compress]{natbib}

\usepackage{tikz}
\definecolor{female}{HTML}{EF8A62}
\definecolor{male}{HTML}{67A9CF}
\definecolor{not-too-happy}{HTML}{FC8D59}
\definecolor{pretty-happy}{HTML}{FFFFBF}
\definecolor{very-happy}{HTML}{91CF60}
\definecolor{grey30}{HTML}{4D4D4D}
\newcommand{\key}[1]
  {\protect \tikz{\fill[#1] rectangle (1ex,1ex);}}

\onlineid{0}
\vgtccategory{Research}
\vgtcinsertpkg

\title{Product plots}
\author{Hadley Wickham, and Heike Hofmann}
\authorfooter{
  \item Hadley Wickham is an Assistant Professor of Statistics at Rice University, Email: hadley@rice.edu.
  \item Heike Hofmann is an Associate Professor of Statistics at Iowa State University.
}
\shortauthortitle{Wickham \MakeLowercase{\textit{et al.}}: Product plots}

\abstract{

}
\keywords{Statistics, joint distribution, conditional distribution, treemap, bar chart, mosaic plot}
\CCScatlist{ % not used in journal version
  \category{H.5.2}{Information Interfaces and Presentation}{ User Interfaces --- Graphical user interfaces (GUI), Interaction styles, Screen design, Evaluation/methodology}
  \CCScat{I.6.8}{Computing Methodologies}%
  {Simulation and Modeling}{Visual Simulation};
}

% \teaser{
% \centering
% \includegraphics[width=16cm]{tx-cancer}
% \caption{One of these plots doesn't belong. These six plots show choropleth maps of cancer deaths in Texas, where darker colors = more deaths. Can you spot which of the six plots is made from a real dataset and not simulated under the null hypothesis of spatial independence?  If so, you've provided formal statistical evidence that deaths from cancer have spatial dependence. See Section~\ref{sec:solutions} for the answer.}
% \label{fig:teaser}
% }

\renewcommand{\manuscriptnotetxt}{\itshape Manuscript received 31 March 2011; accepted XX XXXX XXXX; posted online XX XXXX XXXX; mailed on XX XXXX XXXX.

For information on obtaining reprints of this article, please send email to: tvcg@computer.org.}

\begin{document}
\firstsection{Introduction}
\maketitle
Area plots, like bar charts, mosaic plots and treemaps, are the graphical equivalent of contingency tables, visualising a discrete distribution by mapping proportion to area. This paper presents a unifying framework for area plots, showing how many existing plots are interrelated, and developing a set of tools for creating new plots. In particular, we create a graphical factorisation of area plots by analogy to the mathematical factorisation of joint distributions into products of marginals and conditionals.

@@

\begin{table*}[htb]
  \begin{center}
  \begin{tabular}{llp{9cm}}
    \toprule
    Variable & Description & Values \\
    \midrule
    {\sf age} & age in years & 18--89 \\
    {\sf degree} & highest education & lt high school, high school, junior college, bachelor, graduate \\
    {\sf finrela} & relative financial status & far above average, above average, average, below average, far below average \\
    {\sf happy} & happiness & \key{very-happy} very happy, \key{pretty-happy} pretty happy, \key{not-too-happy} not too happy \\
    {\sf health} & health & excellent, good, fair, poor \\
    {\sf sex} & sex & \key{female} female, \key{male} male\\
    {\sf wtsall} & probability weight & 0.43--6.42 \\
    {\sf year} & year of survey & 1972--2006 \\
    \bottomrule
  \end{tabular}
  \end{center}
  \caption{Description of sample data. Common plot colours are shown next to respective levels.}
  \label{tbl:happy}
\end{table*}

@@

@@

\section{Interaction}
\label{sec:interaction}

@@

\noindent There are 24 possible ways of ordering the variables in the pdf, and each 1d pmf can be displayed in five different ways. For a 4d problem this gives $5 * 24 * (5^3 + 5^4 + 5^2 + 1 + 5^2) = 96120$ plots!  In future work, we hope to connect product plots to statistical models to help guide the choice of appropriate graphic.

% \acknowledgments{ The authors wish to thank Robert Kosara for helpful discussion. This work was partly supported by the National Science Foundation grant DMS0706949. Graphics produced with ggplot2 \citep{me:ggplot2}.}

% bibtool -x graphical-inference -x coordination -o references.bib
\bibliographystyle{ieeetr}
\bibliography{references}

\end{document}
